Make any stupid plot you want

### 2021-04-16 draft/sections/math.tex
compose + equiv sections

\subsubsection{Composition of Artists: +}
\begin{figure}[H]
\begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
    \caption{Each component of a visualization is generated via a different artist.}
    \label{fig:seperate_artists}    
\end{subfigure}
\begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/combined_artist.png}
    \caption{This visualization is the disjoint union of the artists in \autoref{fig:seperate_artists}.}
    \label{fig:combined_artists}
\end{subfigure}
\caption{}
\label{fig:artist_plus}
\end{figure}
To build graphics that are the composites of multiple artists, we define a simple addition operator that is the disjoint union of fiber bundles \dtotal.  For example, a scatter plot $\dtotal_1$ and a line plot $\dtotal_2$ have different \dbase that are mapped to separate \gbase. To fully display both graphics, the composite graphic $\vartist_1 + \vartist_2$ needs to include all records on both $\dbase_1$ and $\dbase_2$, which are the sections on the disjoint union $\dbase_1 \sqcup \dbase_2$. This in turn yields disjoint graphics $\gbase_1 \sqcup \gbase_2$ rendered to the same image. Constraints can be placed on the disjoint union such as that the fiber components need to have the same \vchannel position encodings or that the position \vsection need to be in a specified range. There is a second type of composition where $\dtotal_1$ and $\dtotal_2$ share a base space $\dbase_2 \hookrightarrow \dbase_1$ such that the the artists can be considered to be acting on different components of the same section. This type of composition is important for creating visualizations where elements need to update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
\label{sec:artist_equivalance}
It is impractical to implement an artist for every single graphic; instead we implement an approximation of an the equivalence class of artists $\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}$. Roughly, equivalent artists have the same fiber bundle \vtotal\ and same assembly function \vmark\ but act on different sections \vsection, but we will formalize the definition of the equivalance class in future work. As a first pass for implementation purposes, we identify a minimal \vfiber\ associated with each $\vartist^{\prime}$ that defines what visual characteristics of the graphic must originate in the data such that the graphic is identifiable as a given chart type.
%%
%%\begin{figure}[H]
%%   \caption{Each of these graphics is generated by a different artist \vartist which is the equivalance class of scatter plots $\vartist^{\prime}$\note{this is gonna be a whole bunch of scatter plots}}
%%    \label{fig:artist_equivalence}
%%\end{figure}

For example, a scatter plot of red circles is the output of one artist, a scatter plot of green squares the output of another. These two artists are equivalent since their only difference is in the literal visual encodings (color, shape). Shape and color could also be defined in \vmark\, but the position must come from the fiber $\vfiber=(xpos,ypos)$ since fundementally a scatter plot is the plotting of one position against another\cite{friendlyBriefHistoryData2008}. We also use this criteria to identify derivative types, for example the bubble chart\cite{tufteVisualDisplayQuantitative2001} is a type of scatter where by definition the glyph size is mapped from the data. The criteria for equivalence class membership serves as the basis for evaluating invariance\cite{kindlmann2014algebraic}. 
\end{document}

>>>>

\subsubsection{Composition of Artists: +}
\begin{figure}[H]
\begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/exploding_artist.png}
    \caption{}
    \label{fig:math:seperate}    
\end{subfigure}
\begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/math/combined_artist.png}
    \caption{}
    \label{fig:math:combined}
\end{subfigure}
\caption{Each of the visual elements in \autoref{fig:math:seprate} is generated via a unique artist \vartist. In \autoref{fig:math:seperate}, they are added to the image independent of the other elements, creating an incoherant visualization. In \autoref{fig:math:combined}, these artists are composited before being added to the image. Disjoint union of \dtotal\ aligns the two timeseries with the x and y axis so all these elements use a shared coordinate system. A more complex composition dictates that the legend is connected to the \dtotal\ such that it must use the same color as the data it is identifying.}
\label{fig:artist_plus}
\end{figure}
Visualizations with a single artist do not provide much information, so we define addition operators for generating more complex visualizations. Given the family of artists $(\dtotal_i: i\in I)$ on the same image, the + operator 
\begin{equation}
+ \coloneqq \underset{i\in I}{\sqcup} \dtotal_{i}
\end{equation}
defines a simple composition of artists. For example, the components in \autoref{fig:math:seperate} are each generated by different artists, and a visualization of solely the x axis is rarely all that useful. In \autoref{fig:math:seperate}, these artists are all added to the image independently of the other and therefore there are no constraints on how they are generated in conjunction with each other. In \autoref{fig:math:combined}, the data is joined via disjoint union; doing so aligns the components in \dfiber\ such the \vchannel\ to the same component in \vfiber\ targets the same coordinate system. When artists share a base space $\dbase_2 \hookrightarrow \dbase_1$, a composition operator can be defined such that the artists are acting on different components of the same section. This type of composition is important for visualizations where elements update together in a consistent way, such as multiple views \cite{alboRadarComparativeEvaluation2016a, hullmanKeeping2018} and brush-linked views\cite{beckerBrushingScatterplots1987,bujaInteractiveData1991}.

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
\label{sec:artist_equivalance}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/equivalent_artists.png}
   \caption{Each scatter plot is generated via a unique artist function $\vartist_{i}$, but they only differ in aesthetic styling. Therefore, these artists are all members of an equivalence class $\vartist_{i} \in \vartist{\prime}$}
    \label{fig:math:artist:equivalence}
\end{figure}
Representational invariance, as defined by Kindlmann and Scheidegger, is the notion that visualizations are equivalent if changing the visual representation, such as colors or shapes, does not change the meaning of the visualization\cite{kindlmannAlgebraicProcessVisualization2014}. We propose that visualizations are invariant if they are generated by artists that are members of an equivalence class 
\begin{equation*}
\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}
\end{equation*}
For example, every scatter plot in \autoref{fig:math:artist:equivalence} is a scatter of the same datasets mapped to the \textit{x position} and \textit{y position} in the same way. The scatter plots only differ in the choice of constant visual literals, differing in color and marker shape. Each scatter is generated by an artist $\vartist_{i}$, and every scatter is generated by a member of the equivalence class $\vartist_{i} \in \vartist{\prime}$. Since it is impractical to implement a new artist for every single graphic, the equivalance class provides a way to evaluate an implementation of a generalized artist. Given equivalent, but no necessarily identical, \vchannel, \vmark, and \vindex, two artists are equivalent. This criteria also allows for comparing artists across libraries. 
\end{document}
